Radio Emission from Red-Giant Hot Jupiters

### 2015-12-20 ms.tex
a couple typo fixes

In order to examine the realistic potential to detect radio emission of RGHJs with the current and near-future instruments, we estimate the possible radio emission from known red giants assuming that they have planetary companions. 
For this purpose, we obtained the list of M-type red giants from Table 4 of \citet{dumm1998} with the data of mass, radius, and effective temperature. 
\revise{}{We note that there data do not unambiguously distinguish RG from early AGB stars and thus the list likely includes about 40\% of early AGB stars. }
For each red giant, the mass-loss rate is estimated using the improved Reimers' equation \citep{reimers1975} given by \citet{schroder2005,schroder2007}:
%%%
\begin{equation}
\dot M_\star [M_\odot{\rm /yr}] \sim 8 \times 10^{-14} \frac{\tilde L \tilde R}{\tilde M} \left( \frac{T_{\rm eff}}{4000} \right)^{3.5} \left( 1 + \frac{g_{\odot }}{4300 g_{\star}} \right) \label{eq:mass-loss}
\end{equation}
%%%
where $\tilde L = L_{\star }/L_{\odot }(=4\pi R_{\star }^2 T_{\star }^4)$ is the stellar luminosity, $\tilde R = R_{\star }/R_{\odot }$ is the stellar radius, $\tilde M = M_{\star }/M_{\odot }$ is the stellar mass, and $g$ is the surface gravity. 
The velocity of the stellar wind is assumed to be the escape velocity at $\sim 4 R_{\star }$:
%%%
\begin{equation}
v \sim \sqrt{\frac{2GM_\star}{4R_{\star }}}
\end{equation}
%%%
The age of the system is assumed to be roughly the same as the lifetime of the main-sequence stage, which depends on the stellar mass. We simply assume:
%%%
\begin{equation}
t \sim 10\,{\rm Gyr}\, \left( \frac{M_{\star}}{M_{\odot }} \right)^{-2.5}
\end{equation}
%%%
Combining this relation with equations (\ref{eq:fcyc}) and (\ref{eq:scalingB}) results in
%%%
\begin{equation}
\nu_{\rm cyc,\,max} \sim 20 \,{\rm MHz}\, \left( \frac{M_{\star}}{M_{\odot }} \right)^{1.075} \left( \frac{M_p}{M_{p,\,{\rm J}}} \right)^{1.04} 
\end{equation}
%%%
In addition, the distance to the system is obtained based on the parallax data from Hipparcos datasets\footnote{http://www.rssd.esa.int/index.php?project=HIPPARCOS}.

@@

\revise{}{A more relistic expected number of detection can be also evaluated by combining these results with the probability for a star to have a planet with certain mass and orbit,  %
Observations of main-sequence stars implies that the number of Jovian planets is about a few to ten percent of the number of stars \citep[e.g.][]{cumming2008,cassan2012}. %
Assuming the same population for RG systems, the expected number of actual detection from RGHJs would be order of unity.  
A more detailed examination using the empirical planet mass function by \citet{cumming2008} is given in Appendix \ref{s:expectedvalue}. 
The analysis indicates that if the detection limit is 10$\mu {\rm Jy}$ the total expected value is about 3. }
%However, as we see in the next subsection, the . and the expected value for massive planets ($M_p>5M_{\rm J}$) which supposedly have relatively high cyclotron frequency is about 1-2.

@@

% LOFAR
The LOw-Frequency ARray (LOFAR) operates as two separate arrays defined by their high and low bands, differing configurations, and antenna designs, which are called respectively the Low Band Antennas (LBA) and High Band Antennas (HBA).
LOFAR-LBA operates in the 10-90~MHz range, and LOFAR-HBA operates in the 100-240~MHz range \citep{vanHaarlem2013}.
We consider it as two separate instruments here.
LOFAR-LBA's most sensitive band is centered at 60~MHz, and LOFAR-HBA's is at 150~MHz.
Given the strong frequency dependence of the effective aperture of a dipole antenna, and that the reciever noise temperature for LOFAR $\Trx \sim \Tsky$, we use the LOFAR Image Noise calculator\footnote{Heald; \url{http://www.astron.nl/~heald/test/sens.php}}, which is based on {\it SKA Memo 113}\footnote{Nijboer, Pandey-Pommier, \& de Bruyn; \url{http://www.skatelescope.org/uploaded/59513\_113\_Memo\_Nijboer.pdf}}, to obtain the sensitivity and imaging parameters in Table~\ref{tab:sens}.

@@

\paragraph{Variability of the stellar wind (a few months $\sim $?)}
If the stellar wind passing the planet is variable, any modulation of the plasma density would also lead to time variabililty of planetary radio emission.  Such variability might be expected to occur on timescales of several months to half-a-year as many M giants have semi-regular periodicities on the order of a few hundred days \citep{Kiss:1999aa}.

>>>>

In order to examine the realistic potential to detect radio emission of RGHJs with the current and near-future instruments, we estimate the possible radio emission from known red giants assuming that they have planetary companions. 
For this purpose, we obtained the list of M-type red giants from Table 4 of \citet{dumm1998} with the data of mass, radius, and effective temperature. 
\revise{}{We note that these data do not unambiguously distinguish RG from early AGB stars and thus the list likely includes about 40\% of early AGB stars. }
For each red giant, the mass-loss rate is estimated using the improved Reimers' equation \citep{reimers1975} given by \citet{schroder2005,schroder2007}:
%%%
\begin{equation}
\dot M_\star [M_\odot{\rm /yr}] \sim 8 \times 10^{-14} \frac{\tilde L \tilde R}{\tilde M} \left( \frac{T_{\rm eff}}{4000} \right)^{3.5} \left( 1 + \frac{g_{\odot }}{4300 g_{\star}} \right) \label{eq:mass-loss}
\end{equation}
%%%
where $\tilde L = L_{\star }/L_{\odot }(=4\pi R_{\star }^2 T_{\star }^4)$ is the stellar luminosity, $\tilde R = R_{\star }/R_{\odot }$ is the stellar radius, $\tilde M = M_{\star }/M_{\odot }$ is the stellar mass, and $g$ is the surface gravity. 
The velocity of the stellar wind is assumed to be the escape velocity at $\sim 4 R_{\star }$:
%%%
\begin{equation}
v \sim \sqrt{\frac{2GM_\star}{4R_{\star }}}
\end{equation}
%%%
The age of the system is assumed to be roughly the same as the lifetime of the main-sequence stage, which depends on the stellar mass. We simply assume:
%%%
\begin{equation}
t \sim 10\,{\rm Gyr}\, \left( \frac{M_{\star}}{M_{\odot }} \right)^{-2.5}
\end{equation}
%%%
Combining this relation with equations (\ref{eq:fcyc}) and (\ref{eq:scalingB}) results in
%%%
\begin{equation}
\nu_{\rm cyc,\,max} \sim 20 \,{\rm MHz}\, \left( \frac{M_{\star}}{M_{\odot }} \right)^{1.075} \left( \frac{M_p}{M_{p,\,{\rm J}}} \right)^{1.04} 
\end{equation}
%%%
In addition, the distance to the system is obtained based on the parallax data from Hipparcos datasets\footnote{http://www.rssd.esa.int/index.php?project=HIPPARCOS}.

@@

\revise{}{A more realistic expected number of detection can be also evaluated by combining these results with the probability for a star to have a planet with certain mass and orbit,  %
Observations of main-sequence stars implies that the number of Jovian planets is about a few to ten percent of the number of stars \citep[e.g.][]{cumming2008,cassan2012}. %
Assuming the same population for RG systems, the expected number of actual detection from RGHJs would be order of unity.  
A more detailed examination using the empirical planet mass function by \citet{cumming2008} is given in Appendix \ref{s:expectedvalue}. 
The analysis indicates that if the detection limit is 10$\mu {\rm Jy}$ the total expected value is about 3. }
%However, as we see in the next subsection, the . and the expected value for massive planets ($M_p>5M_{\rm J}$) which supposedly have relatively high cyclotron frequency is about 1-2.

@@

% LOFAR
The LOw-Frequency ARray (LOFAR) operates as two separate arrays defined by their high and low bands, differing configurations, and antenna designs, which are called respectively the Low Band Antennas (LBA) and High Band Antennas (HBA).
LOFAR-LBA operates in the 10-90~MHz range, and LOFAR-HBA operates in the 100-240~MHz range \citep{vanHaarlem2013}.
We consider it as two separate instruments here.
LOFAR-LBA's most sensitive band is centered at 60~MHz, and LOFAR-HBA's is at 150~MHz.
Given the strong frequency dependence of the effective aperture of a dipole antenna, and that the receiver noise temperature for LOFAR $\Trx \sim \Tsky$, we use the LOFAR Image Noise calculator\footnote{Heald; \url{http://www.astron.nl/~heald/test/sens.php}}, which is based on {\it SKA Memo 113}\footnote{Nijboer, Pandey-Pommier, \& de Bruyn; \url{http://www.skatelescope.org/uploaded/59513\_113\_Memo\_Nijboer.pdf}}, to obtain the sensitivity and imaging parameters in Table~\ref{tab:sens}.

@@

\paragraph{Variability of the stellar wind (a few months $\sim $?)}
If the stellar wind passing the planet is variable, any modulation of the plasma density would also lead to time variability of planetary radio emission.  Such variability might be expected to occur on timescales of several months to half-a-year as many M giants have semi-regular periodicities on the order of a few hundred days \citep{Kiss:1999aa}.
